---
output: html_document
editor_options: 
  chunk_output_type: console
---


```{r}
library(purrr)
devtools::load_all()
```


```{r}

seeds <- V(gene42_igraph)[isTerminal]$name

```

## Calculate all shortest paths : stoneTrees method

I'm just looking how all shortest paths are computed in the package.

```{r}
calculateAPSPmodule
```

Compute all shortest path using the same method than the package.

```{r}

allShortestPaths <- all_shortest_paths(gene42_igraph, from = seeds, to = V(gene42_igraph)[name %in% seeds], mode = "all")$res
```

 * How many paths are found ?
 
```{r}
length(allShortestPaths)
```

* From ?

Extract 1st node for each paths.

```{r}
map_chr(allShortestPaths, ~{ .x[1] %>%  as_ids(.) } ) %>% unique
```

That's where the error is. We want all shortest paths from **all** nodes to **all** nodes. It looks like we obtain all shortest paths from the 1st node only.

 * to ?
 
Extract last node for each paths. 

```{r} 
map_chr(allShortestPaths, ~{ .x[length(.x)] %>% as_ids(.) } ) %>% unique
```

That looks alright.

## Calculate all shortest paths : For all nodes

Compute all shortest paths from **all** nodes to all nodes.

```{r}

allShortestPaths2 <-  map(seeds, ~{ all_shortest_paths(gene42_igraph, from = .x, to = V(gene42_igraph)[name %in% seeds], mode = "all")$res } ) %>% unlist(., recursive = FALSE)
```


 * How many paths are found ?
 
```{r}
length(allShortestPaths2)
```

many more paths are found.

* From ?

Extract 1st node for each paths.

```{r}
map_chr(allShortestPaths2, ~{ .x[1] %>%  as_ids(.) } ) %>% unique
```

 * to ?
 
Extract last nodefor each paths. 

```{r} 
map_chr(allShortestPaths2, ~{ .x[length(.x)] %>% as_ids(.) } ) %>% unique
```


## APSP from all to all


```{r}

calculateAPSPmoduleForAll <- function(seeds, searchGraph, omitNA = TRUE){

  searchGraphName <- deparse(substitute(searchGraph)) #Capture search graph name

  validateIsNetwork(searchGraph)

  ifelse(omitNA,
         seeds %<>% na.omit(seeds),
         if(any(is.na(V(searchGraph)$name))) warning("NA nodes found in input seed list but 'omitNA' set to FALSE - this can lead to unstable behaviour") )

  seedsNotInGaph <- setdiff(seeds,V(searchGraph)$name)
  seedsInGaph <- intersect(seeds,V(searchGraph)$name)

  if(length(seedsNotInGaph) == length(seeds)){stop("None of the seeds were found in graph!")}
  if(length(seedsNotInGaph) > 0){ warning("Not all seeds found in graph: ",seedsNotInGaph)}

  APSP_vertices <- map(V(searchGraph)[name %in% seedsInGaph], ~{ 
    
    
    all_shortest_paths(searchGraph,
                       from = .x,
                       to = V(searchGraph)[name %in% seedsInGaph],
                       mode="all") %>% 
      .$res } ) %>% 
    unlist %>% 
    as.integer %>% 
    unique %>% 
    na.omit

  process_APSP_module <- induced_subgraph(searchGraph,V(searchGraph)[APSP_vertices])

  V(process_APSP_module)$isSeed <- FALSE
  V(process_APSP_module)[name %in% seeds]$isSeed <- TRUE

  process_APSP_module %<>% set_graph_attr("SearchNetwork", searchGraphName)

  return(process_APSP_module)
}

```


```{r}
calculateAPSPmoduleForAll(seeds,gene42_igraph)
calculateAPSPmodule(seeds,gene42_igraph)
```

Another example :

```{r}

seeds <- V(lymphomaGraph)$name[V(lymphomaGraph)$nodeScore >0]

calculateAPSPmoduleForAll(seeds,lymphomaGraph)

calculateAPSPmodule(seeds,lymphomaGraph)


```

